---
title: "NY State Cancer Trends"
author: "Devon Morgan"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(readr)
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

```

## Import Data from NY Cancer Registry
This project imports current New York State reported cancer incidence and mortality data reported on health.ny.gov. The data reports cancer incidence and mortality trends by sub-type from 1976-2016. 

```{r}
# Import data from NY Cancer Registry and Convert to Table
url = "https://www.health.ny.gov/statistics/cancer/registry/vol3/v3nys.htm"
cancer_data = read_html(url)

cancer_data %>%
  html_nodes(css = "table")

table_cancer = (cancer_data %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table(fill=TRUE)

# Export data to CSV

write_csv(table_cancer, "./data/cancerdata.csv")

```

## Read in New CSV and Tidy Data

```{r}
clean_cancer_data = read_csv("./data/cancerdata_clean.csv") %>% 
  janitor::clean_names() %>% 
  select(site_of_cancer, diag_death_time_period, incidence_males_rate_per_100_000_males, incidence_females_rate_per_100_000_females)

tidy_male_cancer_data = clean_cancer_data %>% 
  select(-incidence_females_rate_per_100_000_females) %>% 
  filter(site_of_cancer == "Lung and bronchus" | site_of_cancer == "Kidney and renal pelvis" |site_of_cancer =="Urinary bladder (incl. in situ)" | site_of_cancer =="Thyroid" | site_of_cancer =="Non-Hodgkin lymphomas")

tidy_female_cancer_data = clean_cancer_data %>% 
  select(-incidence_males_rate_per_100_000_males) %>% 
  filter(site_of_cancer == "Lung and bronchus" | site_of_cancer == "Kidney and renal pelvis" |site_of_cancer =="Urinary bladder (incl. in situ)" | site_of_cancer =="Thyroid" | site_of_cancer =="Non-Hodgkin lymphomas" |site_of_cancer =="Female breast")
```

## Visualize Trends in Incidence Over Time
The following graphs show 
### Cancer Trends in Males, by Cancer Site
```{r}
ggplot(tidy_male_cancer_data, aes(x = diag_death_time_period, y = incidence_males_rate_per_100_000_males, color = site_of_cancer, group = site_of_cancer)) +
  geom_point(alpha = .5) +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(title = "New York State Cancer Incidence Trends in Males from 1976-2015, by Cancer Site",
      x = "Time Period",
      y = "Incidence Rate per 100,000",
      caption = "Source: New York State Cancer Registry
      Rates are per 100,000 persons, age-adjusted to the 2000 US standard population",
    color = "Cancer Site")

```

### Cancer Trends in Females, by Cancer Site

```{r}
ggplot(tidy_female_cancer_data, aes(x = diag_death_time_period, y = incidence_females_rate_per_100_000_females, color = site_of_cancer, group = site_of_cancer)) +
  geom_point(alpha = .5) +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(title = "New York State Cancer Incidence Trends in Females from 1976-2015, by Cancer Site",
      x = "Time Period",
      y = "Incidence Rate per 100,000",
      caption = "Source: New York State Cancer Registry
      Rates are per 100,000 persons, age-adjusted to the 2000 US standard population",
      color = "Cancer Site"
    )
```

